Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-04 EconGrowth.Rnw
Now creating Cobb-Douglas graph with all energy types.

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  tidy=FALSE,
  comment=NA,
  echo=FALSE,    #Turns off code echo for all chunks. Set TRUE on an individual chunk to see it.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
# List of energy types
energyTypes <- c(Q="Q", X="X", U="U")
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. "1" means first alphabetically.
# The following list sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9)
# Full page lattice plot sizes and other parameters
layoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
figWidth <- 6.5 #Inches
figHeight <- 7.5 #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 0.75 #75% of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)">>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

<<cobb-douglas>>=
####################
# Returns an nls Cobb-Douglas model for the country specified
# Give an energyType ("Q", "X", or "U") if you want to include an energy term
##
cobbDouglasModel <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.

  betaGuess <- 0.3 #if needed
  #Run the non-linear least squares fit to the data.
  if (missing(energyType)){
    # No energy term
    # Establish guess values for alpha and lambda.
    lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
    alphaGuess <- 0.7 #0.7 gives good results for all countries.  
    # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
    modelCD <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha), 
                   data=dataTable, 
                   algorithm = "port",
                   start = list(lambda=lambdaGuess, alpha=alphaGuess),
                   lower = list(lambda=-Inf, alpha=0), #constraints on lambda and alpha
                   upper = list(lambda=Inf, alpha=1))
    return(modelCD)
  } else {
    energyColumnName <- paste("i", energyType, sep="")
    iEnergy <- dataTable[energyColumnName] #grabs the desired energy column
    # Reparameterize to ensure that we meet the constraints:
    # * alpha + beta + gamma = 1.0.
    # * alpha, beta, and gamma are all between 0.0 and 1.0.
    # To do this, we reparameterize as
    # * 0 < a < 1
    # * 0 < b < 1
    # * alpha = min(a, b)
    # * beta = b - a
    # * gamma = 1 - max(a, b)
    # Establish guess values for lambda, alpha, and beta.
    lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
    alphaGuess <- 0.4 
    betaGuess <- 1.0 - alphaGuess
    aGuess <- alphaGuess
    bGuess <- alphaGuess + betaGuess - 0.05
    modelCDe <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^min(a,b) * iLabor^abs(b-a) * iEnergy^(1.0 - max(a,b)), 
                    data = dataTable,
                    algorithm = "port",
                    start = list(lambda=lambdaGuess, a=aGuess, b=bGuess),
                    lower = list(lambda=-Inf, a=0, b=0),
                    upper = list(lambda= Inf, a=1, b=1),
                    control = nls.control(maxiter = 200, 
                                          tol = 1e-05, 
                                          minFactor = 1/1024, 
                                          printEval=TRUE, 
                                          warnOnly=TRUE))
    return(modelCDe)
  }
}
#print(lapply(countryAbbrevs, cobbDouglasModel, energyType="Q"))
model <- cobbDouglasModel(countryAbbrev="ZM", energyType="Q")
print(model)
print(model$convInfo)
#print(lapply(countryAbbrevs, cobbDouglasModel))

@@

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(energyType){
  if (missing(energyType)){
    out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions)) #rbinds all the predictions together
  } else {
    out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions, energyType=energyType))
  }
  colnames(out) <- c("predGDP")
  return(out)
}

@@

############################
# Aggregates the Cobb-Douglas results into a big data table
##
cobbDouglasParamsTable <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas parameters for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}
@

<<CD Params Table, results='asis'>>=
print(cobbDouglasParamsTable(), caption.placement="top", sanitize.colnames.function = identity)
@

<<CD_Params_Graph, fig.width=figWidth, fig.height=3.5, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph <- function(){
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))

  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable, 
                   centers = value, #identifies where the dots should be placed
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(3,1), #3 column, 1 row
                   strip = strip.custom(factor.levels=c(expression(italic(lambda)), #sets labels in strip
                                                        expression(italic(alpha)),
                                                        expression(italic(beta))
                                                        ), 
                                        bg="white"), #sets background color in strip
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   ylim = list(c(-0.05, 0.1), c(-0.5, 1.5), c(-0.5, 1.5)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 y=list(relation="free",                        #allow each axis to be different
                                        at=list(c(0.0, 0.05, 0.1),   #tick mark locations on y axis
                                                c(0.0, 0.5, 1.0), 
                                                c(0.0, 0.5, 1.0)
                                                )
                                        )
                                 )
                   )
  return(graph)
}
createCDParamsGraph()
@

<<CD_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas (without energy) results.">>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the models
  dataTable <- cbind(dataTable, predictions) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1) #line widths. 0 means no line.
  colors <- c("black", "black") #line and symbol colors
  symbols <- c(1, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "Predicted")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,8), #y axis limits
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@

@@

  print(countryName)
  energyColumnName  <- paste("i", energyType, sep="")
  #print(energyColumnName)

@@

  summaryCDe <- summary(modelCDe) # Gives the nls summary table
  print(summaryCDe)

@@

  a <- as.numeric(coef(modelCDe)["a"])
  b <- as.numeric(coef(modelCDe)["b"])
  print(a)
  print(b)
  lambda <- as.numeric(coef(modelCDe)["lambda"])
  alpha <- a
  beta <- b - a
  gamma <- 1.0 - alpha - beta

>>>>

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(tikzDevice) #See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  echo=FALSE,    #Turns off code echo for all chunks. Set TRUE on an individual chunk to see it.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
# List of energy types
energyTypes <- c(Q="Q", X="X", U="U")
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. "1" means first alphabetically.
# The following list sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9)
# Full page lattice plot sizes and other parameters
layoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
figWidth <- 6.5 #Inches
figHeight <- 7.5 #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 1.0 #0.75 #75% of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", eval=FALSE>>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

<<cobb-douglas>>=
####################
# Returns an nls Cobb-Douglas model for the country specified
# Give an energyType ("Q", "X", or "U") if you want to include an energy term
##
cobbDouglasModel <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  #Run the non-linear least squares fit to the data based on whether we have an energy type or not.
  if (missing(energyType)){
    # No energy term desired in the Cobb-Douglas equation.
    # Establish guess values for alpha and lambda.
    lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
    alphaGuess <- 0.7 #0.7 gives good results for all countries.  
    # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
    modelCD <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha), 
                   data=dataTable, 
                   start = list(lambda=lambdaGuess, alpha=alphaGuess),
                   #Include the next 3 lines to fit with constraints.
                   #algorithm = "port",
                   #lower = list(lambda=-Inf, alpha=0), #constraints on lambda and alpha
                   #upper = list(lambda=Inf, alpha=1)
                   )
    return(modelCD)
  } else {
    # We need to do the Cobb-Douglas fit with the desired energy data.
    # To achieve the correct fit, we'll change the name of the desired column
    # to "iEToFit" and use "iEToFit" in the nls function. 
    energyColumnName <- paste("i", energyType, sep="") #Energy term desired in the Cobb-Douglas equation.
    eColIndex <- which(names(dataTable) %in% energyColumnName) #Find index of desired column
    colnames(dataTable)[eColIndex] <- "iEToFit" #Change desired column name to "iEToFit"
    # Reparameterize to ensure that we meet the constraints:
    # * alpha + beta + gamma = 1.0.
    # * alpha, beta, and gamma are all between 0.0 and 1.0.
    # To do this, we reparameterize as
    # * 0 < a < 1
    # * 0 < b < 1
    # * alpha = min(a, b)
    # * beta = b - a
    # * gamma = 1 - max(a, b)
    # Establish guess values for lambda, alpha, and beta.
    lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
    alphaGuess <- 0.9
    betaGuess <- 1.0 - alphaGuess
    aGuess <- alphaGuess
    bGuess <- alphaGuess + betaGuess - 0.05
    # Now actually do the fit, using the column name "iEToFit".
    modelCDe <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^min(a,b) * iLabor^abs(b-a) * iEToFit^(1.0 - max(a,b)), 
                    data = dataTable,
                    start = list(lambda=lambdaGuess, a=aGuess, b=bGuess),
                    control = nls.control(maxiter = 200, 
                                          tol = 1e-05, 
                                          minFactor = 1/4096, 
                                          printEval=FALSE, #Tells whether to print details of curve fit process.
                                          warnOnly=TRUE),
                    #Include the next 3 lines to fit with constraints.
                    #algorithm = "port",
                    #lower = list(lambda=-Inf, a=0, b=0),
                    #upper = list(lambda= Inf, a=1, b=1)
                    )
    return(modelCDe)
  }
}

@@

print("US"); print(cobbDouglasPredictions("US", "Q"))
print("UK"); print(cobbDouglasPredictions("UK", "Q"))
print("JP"); print(cobbDouglasPredictions("JP", "Q"))
print("CN"); print(cobbDouglasPredictions("CN", "Q"))
print("ZA"); print(cobbDouglasPredictions("ZA", "Q"))
print("SA"); print(cobbDouglasPredictions("SA", "Q"))
print("IR"); print(cobbDouglasPredictions("IR", "Q"))
print("TZ"); print(cobbDouglasPredictions("TZ", "Q"))
print("ZM"); print(cobbDouglasPredictions("ZM", "Q"))

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(energyType){
  if (missing(energyType)){
    out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions)) #rbinds all the predictions together
    colnames(out) <- c("predGDP")
  } else {
    out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions, energyType=energyType))
    colnames(out) <- c(paste("predGDP", energyType, sep=""))
    print(colnames(out))
  }
  return(out)
}

@@

############################
# Aggregates the Cobb-Douglas results into a big data table
##
cobbDouglasParamsTableNoEnergy <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas (without energy) for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}
@

<<CD Params Table, results='asis'>>=
print(cobbDouglasParamsTableNoEnergy(), caption.placement="top", sanitize.colnames.function = identity)
@

<<CD_Params_Graph, fig.width=figWidth, fig.height=3.5, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=FALSE>>=
createCDParamsGraph <- function(){
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))

  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable, 
                   centers = value, #identifies where the dots should be placed
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(3,1), #3 column, 1 row
                   strip = strip.custom(factor.levels=c("$\\lambda$", "$\\alpha$", "$\\beta$"),  #sets labels in strip 
                                        bg="white"), #sets background color in strip
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   ylim = list(c(-0.05, 0.1), c(-0.5, 1.5), c(-0.5, 1.5)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.0, 0.05, 0.1),   #y tick mark for lambda
                                                c(0.0, 0.5, 1.0), 
                                                c(0.0, 0.5, 1.0)
                                                )
                                        )
                                 )
                   )
  return(graph)
}
createCDParamsGraph()
@

<<CD_GDP_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas results.">>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the CD model without energy
  predictionsQ <- cobbDouglasPredictionsColumn("Q") #Predictions with Q
  predictionsX <- cobbDouglasPredictionsColumn("X") #Predictions with X
  dataTable <- cbind(dataTable, predictions, predictionsQ, predictionsX) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "red", "blue") #line and symbol colors
  symbols <- c(1, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP+predGDPQ+predGDPX ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "No energy", "With q", "With x")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allow each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
#                  ylim=c(0.5,3), #y axis limits
                  ylim=list(c(1,8), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3)),

                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@

@@

  #print(countryName)
  energyColumnName  <- paste("i", energyType, sep="")
  #print(energyColumnName)

@@

  summaryCDe <- summary(modelCDe) # Gives the nls summary table
  #print(summaryCDe)

@@

  a <- as.numeric(coef(modelCDe)["a"])
  b <- as.numeric(coef(modelCDe)["b"])
  #print(a)
  #print(b)
  lambda <- as.numeric(coef(modelCDe)["lambda"])
  alpha <- a
  beta <- b - a
  gamma <- 1.0 - alpha - beta
